$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-05-11 paper/e6ssm-tuning-paper.tex
added refs for two-loop Higgs and tried to address referees point 4 by modifying wording.

Using the approach outlined above, we are able to scan the low energy
parameter space of the MSSM and E$_6$SSM and calculate the fine tuning
in each.  To do so, we implemented the above expressions for computing
the fine tuning in a modified version of the E$_6$SSM spectrum
generator that was used in \cite{Athron:2013ipa}.  This code
implemented two-loop RGEs for all parameters except the soft scalar
masses.  In order to properly include the fine tuning impact of the
$SU(3)$ gaugino soft mass $M_3$, we have extended the original code to
make use of the two-loop RGEs generated by SARAH
\cite{Staub:2009bi, Staub:2010jh, Staub:2012pb, Staub:2013tta} and
FlexibleSUSY \cite{Athron:2014yba}, which also makes use of SOFTSUSY
\cite{Allanach:2001kg, Allanach:2013kza}.  The CP-even Higgs masses are
calculated including the leading one-loop effective potential
contributions given in \cite{Athron:2009bs} and the leading two-loop
contributions.  To scan over the MSSM parameter space, the equivalent
MSSM fine tuning expressions were implemented into a modified version
of SOFTSUSY 3.3.10 \cite{Allanach:2001kg}.  For consistency with the results
produced in the $E_6$ models, and for computational speed, for our main scans
only the dominant one- and two-loop corrections to the CP-even Higgs masses
were included.  Finally, in all of the results below the fine tuning was
evaluated at the scale $Q = M_{\textrm{SUSY}} =
\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$, where $m_{\tilde{t}_{1,2}}$ are the
running $\overline{\textrm{DR}}$ stop masses evaluated at
$Q = M_{\textrm{SUSY}}$.

@@

To illustrate this, in the left panel of Fig.~\ref{Fig:mssmstoptuning}
we show the variation in fine tuning for $M_X = 20$ TeV, $50$ TeV, $100$
TeV and $10^{16}$ GeV when we scan over the stop masses and mixing,
with $500 \textrm{ GeV} \leq m_{Q_3} , m_{u_3} \leq 10 \textrm{ TeV}$ and
$-3810\textrm{ GeV} \leq A_t \leq -20 \textrm{ GeV}$.  The remaining
parameters we fix such that at $M_{\rm SUSY}$ they have the values
$\mu = -97.5$ GeV, $B = -84.8$ GeV, $M_1 = 92.1$ GeV, $M_2 = 95.9$ GeV,
$M_3 = 352$ GeV, $A_b = -117.9$ GeV, $A_\tau = -7.8$ GeV, $m_{L_i} = 400$ GeV,
$m_{e_i} = 204$ GeV, $m_{Q_{1,2}} = 438$ GeV, $m_{u_{1,2}} = 436$ GeV and
$m_{d_i} = 438$ GeV ($i = 1 , 2 , 3$).  Here we denote by $M_1$, $M_2$ and
$M_3$ the soft gaugino masses for $U(1)$, $SU(2)$ and $SU(3)$, while $A_b$ and
$A_\tau$ are soft trilinear couplings and the $m_\phi$ are soft scalar
masses for the indicated fields.  The soft bilinear coupling $B$ is defined
such that at tree level the mass of the CP-odd MSSM Higgs boson reads
$m_A^2 = 2 B \mu / \sin 2\beta$.  All off-diagonal couplings and scalar masses
are set to zero, as are the first and second generation Yukawa couplings
and soft trilinears.  Although we should stress that making
this choice will lead to a spectrum which is in conflict with the LHC
limits, doing so ensures that fine tuning due to the other parameters
is small, so that we avoid washing out the fine tuning impact of the
stops when the tuning is small\footnote{For models in which the
spectrum is heavier, when the stop masses are small the fine tuning
reaches a lower bound imposed by other heavier parameters.} as can
be the case when the stop masses are less than 1 TeV.  Note that the
Higgs mass is also allowed to vary in this scan, as shown in the right
panel of Fig.~\ref{Fig:mssmstoptuning}.  This illustrates the tuning
problem which people have been worrying about since the discovery of
the 125 GeV Higgs boson as we see that raising the stop masses is also
pushing up the Higgs mass, meaning that heavier Higgs masses require
more fine tuning.  However for a low value of the UV scale this tuning
is not so severe unless the stops are very heavy, and a 125 GeV Higgs
can be obtained without much tuning in this unrealistic case where we
have minimized other sources of tuning.  On the other hand the tuning
becomes more severe as we increase the cutoff such that for $M_X =
10^{16}$ GeV a lightest stop masses of $1-3$ TeV can result in a fine
tuning of $\approx 100 - 1000$ and the minimum tuning we see
consistent with a $125$ GeV Higgs is $\approx 200$, as shown in
Fig.~\ref{Fig:mssmstoptuning}.

>>>>

Using the approach outlined above, we are able to scan the low energy
parameter space of the MSSM and E$_6$SSM and calculate the fine tuning
in each.  To do so, we implemented the above expressions for computing
the fine tuning in a modified version of the E$_6$SSM spectrum
generator that was used in \cite{Athron:2013ipa}.  This code
implemented two-loop RGEs for all parameters except the soft scalar
masses.  In order to properly include the fine tuning impact of the
$SU(3)$ gaugino soft mass $M_3$, we have extended the original code to
make use of the two-loop RGEs generated by SARAH \cite{Staub:2009bi,
  Staub:2010jh, Staub:2012pb, Staub:2013tta} and FlexibleSUSY
\cite{Athron:2014yba}, which also makes use of SOFTSUSY
\cite{Allanach:2001kg, Allanach:2013kza}.  The CP-even Higgs masses
are calculated including the leading one-loop effective potential
contributions given in \cite{Athron:2009bs} and the leading two-loop
contributions from \cite{King:2005jy} which are a generalsiation of
the correctsions in the MSSM and NMSSM calculated using effective
field theory techniques \cite{Carena:1995wu,Ellwanger:1999ji}.  To
scan over the MSSM parameter space, the equivalent MSSM fine tuning
expressions were implemented into a modified version of SOFTSUSY
3.3.10 \cite{Allanach:2001kg}.  For consistency with the results
produced in the $E_6$ models, and for computational speed, for our
main scans only the dominant one- and two-loop corrections to the
CP-even Higgs masses were included.  Finally, in all of the results
below the fine tuning was evaluated at the scale $Q =
M_{\textrm{SUSY}} = \sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$, where
$m_{\tilde{t}_{1,2}}$ are the running $\overline{\textrm{DR}}$ stop
masses evaluated at $Q = M_{\textrm{SUSY}}$.

@@

To illustrate this, in the left panel of Fig.~\ref{Fig:mssmstoptuning}
we show the variation in fine tuning for $M_X = 20$ TeV, $50$ TeV, $100$
TeV and $10^{16}$ GeV when we scan over the stop masses and mixing,
with $500 \textrm{ GeV} \leq m_{Q_3} , m_{u_3} \leq 10 \textrm{ TeV}$ and
$-3810\textrm{ GeV} \leq A_t \leq -20 \textrm{ GeV}$.  The remaining
parameters we fix such that at $M_{\rm SUSY}$ they have the values
$\mu = -97.5$ GeV, $B = -84.8$ GeV, $M_1 = 92.1$ GeV, $M_2 = 95.9$ GeV,
$M_3 = 352$ GeV, $A_b = -117.9$ GeV, $A_\tau = -7.8$ GeV, $m_{L_i} = 400$ GeV,
$m_{e_i} = 204$ GeV, $m_{Q_{1,2}} = 438$ GeV, $m_{u_{1,2}} = 436$ GeV and
$m_{d_i} = 438$ GeV ($i = 1 , 2 , 3$).  Here we denote by $M_1$, $M_2$ and
$M_3$ the soft gaugino masses for $U(1)$, $SU(2)$ and $SU(3)$, while $A_b$ and
$A_\tau$ are soft trilinear couplings and the $m_\phi$ are soft scalar
masses for the indicated fields.  The soft bilinear coupling $B$ is defined
such that at tree level the mass of the CP-odd MSSM Higgs boson reads
$m_A^2 = 2 B \mu / \sin 2\beta$.  All off-diagonal couplings and scalar masses
are set to zero, as are the first and second generation Yukawa couplings
and soft trilinears.  Although we should stress that making
this choice will lead to a spectrum which is in conflict with the LHC
limits, doing so ensures that fine tuning due to the other parameters
is small, so that we avoid washing out the fine tuning impact of the
stops when the tuning is small\footnote{For models in which the
spectrum is heavier, when the stop masses are small the fine tuning
reaches a lower bound imposed by other heavier parameters.} as can
be the case when the stop masses are less than 1 TeV.  Note that the
Higgs mass is also allowed to vary in this scan, as shown in the right
panel of Fig.~\ref{Fig:mssmstoptuning}.  This illustrates the tuning
problem which people have been worrying about since the discovery of
the 125 GeV Higgs boson as we see that raising the stop masses is also
pushing up the Higgs mass, meaning that heavier Higgs masses require
more fine tuning.  However for a low value of the UV scale this tuning
is not so severe unless the stops are very heavy, and a 125 GeV Higgs
can be obtained without much tuning in this unrealistic case where we
have minimized other sources of tuning.  On the other hand the tuning
becomes more severe as we increase the cutoff such that for $M_X =
10^{16}$ GeV a lightest stop masses of $1-3$ TeV can result in a fine
tuning of $\approx 100 - 1000$ and the minimum tuning we find \footnote{Note that in the calculation of the Higgs mass there is a significant theoretical error even with leading two loop corrections, which should be considered when interpreting what such results for the miniumum fine tuining in the model, given the discovery of a 125 GeV Higgs.}
for a $125$ GeV Higgs is $\approx 200$, as shown in
Fig.~\ref{Fig:mssmstoptuning}.
